VOI-based Monte Carlo Sampling

### 2011-09-14 paper/voisam/paper.tex
first draft for the bounds

\begin{abstract}
\end{abstract}

@@

@@

@@

>>>>

\begin{abstract}
Upper bounds for the VOI are provided for pure exploitaiton in the
Mulit-armed Bandit Problem. Sampling policies based on the upper
bounds are suggested.
\end{abstract}

@@

The intrinsic VOI of pulling an arm is the expected decrease
in the regret compared to selecting an arm without pulling any arm at
all. The intrinsic VOI can be estimated as the intrinsic 
value of perfect information $\Lambda_i$ about the mean reward of the $i$th arm. Two
cases are possible:
\begin{itemize}
\item the arm ($\alpha$) with the highest sample mean is pulled, and the 
mean of the arm is lower than the sample mean the second-best arm ($\beta$);
\item another arm is pulled, and the mean of the arm is higher
than the current highest sample mean ($\alpha$).
\end{itemize}
\begin{eqnarray}
\Lambda_\alpha&=&\int_0^{\overline X_\beta}p(\IE[X_\alpha]=x)(\overline X_\beta-x)dx\nonumber\\
\Lambda_{i|i\ne\alpha}&=&\int_{\overline X_\alpha}^1p(\IE[X_i]=x)(x-\overline X_\alpha)dx
\end{eqnarray}
Using the concentration inequalities, $\Lambda_i$ can be bounded from
above as the probability that a different arm is selected times the
maximum possible increase in the reward:
\begin{eqnarray}
\Lambda_\alpha&\le&\Pr(\IE[X_\alpha]\le\overline X_\beta)\overline X_\beta\nonumber\\
\Lambda_{i|i\ne \alpha}&\le&\Pr(\IE[X_i]\ge\overline
X_\alpha)(1-\overline X_\alpha)
\label{eq:lambda-general-bounds}
\end{eqnarray}

The probabilities in bounds~(\ref{eq:lambda-general-bounds}) can be
bounded from above by the concentration
inequalities~(\ref{eq:conc-hoeffding}, \ref{eq:conc-empbernstein});
thus upper bounds on $\Lambda$ corresponding to each of the
concentration inequalities are obtained. For the Hoeffding's
inequality~(\ref{eq:conc-hoeffding}):
\begin{eqnarray}
\Lambda_\alpha&\le&\hat\Lambda_\alpha^h=\overline X_\beta \exp\left(-2n(\overline  X_\alpha - \overline X_\beta)^2\right)\nonumber\\
\Lambda_{i|i\ne \alpha}&\le&\hat\Lambda_{i|i\ne \alpha}^h=(1-\overline X_\alpha) \exp\left(-2n(\overline X_\alpha - \overline X_i)^2\right)
\label{eq:lambda-hoeffding-bounds}
\end{eqnarray}
For the empirical Bernstein's inequality~(\ref{eq:conc-empbernstein}):
\begin{eqnarray}
\Lambda_\alpha&\le&\hat\Lambda_\alpha^h=\overline X_\beta \exp\left(-\frac {n(\overline X_\alpha - \overline X_\beta)^2} {10(\overline X_\alpha - \overline X_\beta)+2\overline\sigma_{\alpha,n}^2}\right)\nonumber\\
\Lambda_\alpha&\le&\hat\Lambda_{i|i\ne \alpha}^h=(1-\overline X_\alpha) \exp\left(-\frac {2n(\overline X_\alpha - \overline X_i)^2} {10(\overline X_\alpha - \overline X_i)+2\overline\sigma_{i,n}^2}\right)
\label{eq:lambda-bernstein-bounds}
\end{eqnarray}
and a tighter bound can be constructed as $\hat\Lambda^{h,b}=\min(\hat\Lambda^h,\hat\Lambda^b)$.

@@

Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
Lambda$ on the perfect value of information should be pulled at each
step. This way, arms known to have a low VOI  would be pulled less frequently.

@@

The upper bounds~(\ref{eq:lambda-hoeffding-bounds}, \ref{eq:lambda-bernstein-bounds}) decrease exponentially with the
number of pulls $n$. When the upper bound of the VOI for all arms
becomes lower than a threshold $\lambda$, which can be chosen based on
resource constraints, the sampling should be stopped and an arm should
be chosen.
